---
title: "A3: Incarceration"
output: html_document
authors: Vatsal Chandel
---

### Introduction

- This is an analysis report on incarcerated individuals. Specifically, it is data on prison population per county and state. 

- Some of the variables that I focus on are the total populations in the counties (`total_pop`), the prison populations (`total_prison_pop`), the years (`year`), the different counties (`county_name`), states (`state`), genders and race.

- I chose to look at these variables in particular because I wanted to see how the incarceration data compared between different genders and races and over time. I was also curious about the incarceration data looked in Washington. 

### Summary Information
```{r, message = FALSE, echo = FALSE, code = readLines("summary.R")}

```


- The number of prison reports based on `urbanicity` that came in 2018 in the United States were `r count_pop_2018`
  - Of that, `r sum_rural` were rural, `r sum_small_mid` were small/mid, `r sum_suburban` were suburban, and `r sum_urban` were urban 

- Focusing specifically on Washington in 2000, the county that had the highest prison population was `r county_wa_highest_prison_pop`, and had `r wa_highest_prison_pop` individuals.

- In the United States, the county that had the highest prison population was `r county_highest_prison_pop`, with `r highest_prison_pop` individuals in 2000. 

- Looking at the different races in between 2000 and 2016 in King County:
  - Asian or Pacific Islander had a change of `r aapi_over_time`. This means that over time, the number of Asian or Pacific Islanders decreased. 
  - Black had a change of `r black_over_time`. This means that over time, the number of Black indivudals decreased. 
  - Latinx had a change of `r latinx_over_time`. This means that over time, the number of Latinx individuals decreased. 
  - Native had a change of `r native_over_time`. This means that over time the number of Native individuals decreased. 
  - Other races had a change of `r other_over_time`. This means that over time the number of Other race individuals increased. 
  - White had a change of `r white_over_time`. This means that over time the number of White individuals decreased. 
  


### The Dataset

Who collected the data?  

- The set that I used was the data set about prison population per county/state. This data was collected by the Vera Institute of Justice 

How was the data collected or generated?  

- This data was assembled using data collected by the U.S Department of Justice Bureau of Justice Statistics. It was also combined with other data set to fill in some of the empty spaces. 

Why was the data collected?  

- The data was collected to shine more light on the causes and consequences of who all are being sent to prison or jail. Making the data more revolved around county data makes studies more grounded and understandable.


How many observations (rows) are in your data?

- I primarily used the data on prison population per county/state which had `r rows_of_data` rows.  

How many features (columns) are in the data?

- I primarily used the data on prison population per county/state which had `r cols_of_data` columns. 

What, if any, ethical questions or questions of power do you need to consider when working with this data?

- Some questions to consider when working with this data is that it can be considered sensitive information. Lots of individuals have been sent to prison/jail and it can be a huge thing in their lives which means analysis on this data should not be taken lightly.  

What are possible limitations or problems with this data?   (at least 200 words)

- One limitation or problem with this data set is that it does not contain much information about other gender identities. It contains male and female, but has no information on individuals who do not identify as those. This could make analysis on other genders and identities much harder. 
- Another problem with this data set has to do with missing values. For example, in the `aapi_pop_15to64` there are about 153,811 rows. Out of those about 62,780 are missing. This means that about 41% of the data in this column are missing, which could make data analysis much more difficult. 
- Another problem with this data set is that it does not contain clear information on age. It contains columns such as `total_pop_15to64` which tells us how many people are in between the ages of 15 and 64. However, it does not give us information on a specific age. It also does not give us a clear idea on the number of individuals who are older than 64 or younger than 15.
- Another problem with this data set is that it does not contain information about what `other_race_prison_pop` means. The documentation talks about how it is other or unknown racial categories but that does not give much information to work on. This also heavily limits individuals because if they are not Asian or pacific islander, black, latinx, native american, or white, then they have to be classified as 'other'
- Also looking at the data, it becomes clear that there is not much data on the 2018 and 2017 year. This makes most recent analysis difficult because the next best year is 2016 which is fairly old. 

### Trends Over Time Chart

Chart: 

```{r, message = FALSE, echo = FALSE, fig.height = 5, fig.width= 10, fig.align='center', code = readLines("chart1.R")}
```

Description and Why: 

- This line graph shows how many of the different races (Asian or Pacific Islander, Black, Latinx, Native American, Other/Unknown, and White) in the data set went to Prison in King County from 2000 to 2016. I chose this range because it contains the most complete data.
- I included this graph because it shows the difference in races and how their populations changed over time in that prison.
- One trend that shows up is that Black, Latinx, Native American, White, and Asian or Pacific Islander tend to decrease in population while Other/Unknown went up. 



### Variable Comparison Chart

Chart: 
```{r, message = FALSE, echo = FALSE, fig.height = 5, fig.width= 10, fig.align='center', code = readLines("chart2.r")}
```


Description and Why: 

- This line graph compares the Male and Female populations in prison from 2000 to 2016 in King County. I chose this range because it contains the most complete data. 
- I included this graph because it compares the different Male and Female population values. 
- One trend that shows up is that the number of men in prison is much higher than women. It is also clear that the both populations are decreasing as time goes on. 

### Map

Chart:

```{r, message = FALSE, echo = FALSE, fig.height = 5, fig.width= 10, fig.align='center', code = readLines("map.r")}
```

Description and Why: 

- This is a map of all of Washington's county in the year 2000 and how the prisons total populations compare.

- I chose the year 2000 because the data was complete and it was interesting to see data in this year. 

- This was included to show which counties have the highest number of prisoners in Washington. One pattern visable from the map is that King County and Pierce County have the highest population. This tends to make sense because they have some of the highest populations in Washington. 



